SAT Competition 2020\tnoteref{title}

### 2021-06-16 paper/main.tex
explain no proof rule for Parallel

The success of SAT would not be possible without the persistent efforts of the SAT community to further improve the performance and robustness
of SAT solvers. The SAT Competition series, with a history dating back to early 90s, aims to support and provide further incentives 
for maintaining this progress. Organized yearly as an international open event, SAT Competitions (and their variants
in the forms of SAT Races and a SAT Challenge)~\cite{DBLP:journals/amai/SimonBH05,DBLP:conf/sat/BerreS03,DBLP:conf/sat/BerreS04,DBLP:journals/aim/JarvisaloBRS12,DBLP:journals/ai/BalintBJS15,DBLP:conf/aaai/BalyoHJ17,DBLP:journals/jsat/HeuleJS19}
 have a consistent track record in receiving tens of solver submissions yearly, submitted
by the community at large for obtaining a snapshot of the current state-of-the-art in practical SAT solving. Alongside participating solvers,
the competition invites through open calls submissions of benchmark instances representing, in particular, new interesting applications 
scenarios of SAT solvers. Indeed, in addition to evaluating recently developed solvers, an important aspect of the SAT competition series is
to collect on a yearly basis new benchmark sets, consisting of instances from various different application settings, which together with benchmark sets from previous years constitute a standard dataset for use in research papers and SAT solver development.

@@

However, solvers in No-Limits still competed against all other solvers submitted to the Main Track.
Thus, to deserve a mention, a No-Limits solver would need rank among the best-performing solvers among the Main Track participants.
The top ranked solvers in the No-Limits track were the same as in the Main track. 
This also indicates the stability of results under the exclusion of old benchmark instances.

Complementing the generality advocated by the standard SAT Competition tracks, in which solvers are evaluated
on a set of benchmarks including instances from various types of different problem domains, for 2020 
the organizers aimed to experiment with the potential of a more application-specific track,
each year highlighting a different problem domain where the SAT solving technology helps to advance the state of the art.
In 2020, the Planning track represented the first trial instantiation of this idea.
The focus of this track was specifically on efficiently solving
instances arising from the domain of SAT-based automated planning.\textcolor{red}{Add references} 
Automated planning was chosen as the target problem domain of this first instantiation of the domain-specific tracks
due to its centrality as oneof the first breakthrough applications of SAT solvers.
To this end, solvers participating in the Planning
track were evaluated on 200 benchmark instances encoding planning problems.
The same rules for participation as in the Main track applied to the Planning track.
Solvers submitted to the Main track automatically participated in the Planning track.

The traditional Hack track (est. 2009 as \solver{Minisat} Hack track) was organized as a sub-track of the Main track for hacks of \solver{Glucose 3}\textcolor{red}{Add reference}. 
In the past, several advances in SAT solving required only small modifications of an established solver to achieve a considerable contribution. 
Hack tracks encourage participation of such small modifications. 
The limit for being considered a ``hack'' was---somewhat arbitrarily\footnote{The specific threshold for edit distance is not central here; the idea is essentially to only allow relatively small changes to the Glucose
code base, i.e., ``quick hacks'' to Glucose.}---set to 1000 non-space character edit distance from the sources of \solver{Glucose 3}. 
Unfortunately, in 2020 there were not enough participants in this sub-track such that we do not report on it in the results section.

We evaluated all 64 solver submissions (including different configurations of specific solvers) 
to the Main track. Out of the 64 solvers, eight were explicitly submitted to the No-Limits track. Four solvers
 were demoted to the No-Limits track due to outputting invalid unsatisfiability proof certificates. 
Six solvers were disqualified due to outputting truth assignments which did not satisfy the corresponding benchmark instance.
This left us with 46 configurations of 22 solvers, inluding only one Glucose hack.

@@

In the Incremental Library track the underlying idea is to mimic scenarios
where a SAT solver is used as a back-end solver in a more complex tool
(typically solving a harder problem than SAT) and is called multiple times before 
the enclosing tool reaches its final state. 
``Incremental'' here refers to the idea that the individual calls to the SAT solver are not independent, but may share a common subset of the input clauses or differ in the presence of additional unit clause assumptions~\cite{Nadel:2014:Incremental,Fazekas:2019:IncrementalInprocessing}.\textcolor{red}{Should we not cite the 2004 temporal induction paper by Niklases here?}
Examples for applications of incremental SAT solving are counterexample-guided eastraction refinement (CEGAR) based approaches, e.g., for Bounded Model Checking~\cite{Strichman:2005:ARBMC}, SAT-based planning~\cite{Balyo:2017:IncrPlanning} and satisfiability modulo theories (SMT) solvers~\cite{Brummayer:2009:Boolector}.

Instead of using or extending the DIMACS input format, in the the Incremental Library track
a general incremental interface called IPASIR (Re-entrant Incremental Solver API) 
is employed~\cite{Balyo:2015:SATRace}. The idea is that we actually run the 
enclosing tool on its own benchmark and communicate with the competing SAT solver 
through this interface API. 
%
SAT solvers submitted to this track need hence implement the require interface.
Furthermore, it should be noted that the solutions output by a solver may, in general, influence
the fortchoming invocations of the solver.
% E.g. Smaller unsat cores may allow the enclosing tool to converge using fewer SAT calls!

@@

The Parallel track evaluates the runtime performance of SAT solvers making use of multiple processor cores
in terms of wall-clock time.
The benchmarks %and the timeout (5000 seconds) 
are the same as in the Main track. 
%\textcolor{red}{repetitive? also mentioned below?}. 
%Each solver in this track was run on a m4.16xlarge machine from Amazon Web Services (AWS).
%These machines have 64 virtual cores and 256GB of memory. 
In contrast to the Main track, proof logging for unsatisfiable instances is not required in the Parallel track.

@@

For data-driven selection and distribution of benchmark instances, we use GBD~Tools\footnote{\url{https://pypi.org/project/gbd-tools/}} which facilitates querying for instances with desired properties, e.g., instance author, family, result or solver-runtime~\cite{Iser:2018:GBD}. 
We also run GBD~Tools on a server to distribute benchmark instances and attribute databases.\footnote{\url{https://gbd.iti.kit.edu}}

@@

The ``Bring Your Own Benchmarks'' (BYOB) rule, first  established in SAT Competition 2017~\cite{SC2017}, was again followed
in 2020.
By this rule, solver authors are required to submit 20~new benchmark instances in order to participate in the competition. At least 10~of these instances are required to be ``interesting'', meaning that \solver{Minisat} must not solve it in less than one minute.

As can be seen in Table~\ref{tab:families}, 27~authors contributed a set of 1260~previously unseen benchmark instances from a wide range of different instance families. 
In order to create a balanced set of 300 new benchmarks and 100 old benchmarks, i.e., instances from previous competitions, we executed the following steps of \emph{rule-based randomization}.

To compile the set of 300 new instances, we first applied a hardness criterion by filtering out all instances solved by \solver{Minisat} in less than 10 minutes. 
From the resulting 1012 instances, we randomly selected $k$~instances per author, with a minimum $k$~such that the result set contains at least 300 instances.
Algorithm~\ref{algo:select} depicts the details. 
We randomly selected 7~satisfiable and 7~unsatisfiable instances per author (Lines~3 and~4) and added instances of yet unknown result if this did not yield a total of 14~instances (Lines~5--7). 
Of the such obtained 308~instances, we randomly removed 8~satisfiable instances, yielding a total of 114~satisfiable, 78~unsatisfiable and 108~instances of unknown result.

We augmented this set with 100 instances from previous competitions as follows. 
In order to further balance the amount of satisfiable and unsatisfiable instances in the new benchmark set, we randomly selected 21~satisfiable, 57~unsatisfiable and 22~unknown instances. 
With additional constraints, we made sure not to select instances of families which are already represented in the set of~300 new instances (cf. Table~\ref{tab:families}). 
We also excluded random, agile and planning instances (due to the Planning track). 
The final selection of benchmark instances contains 135~satisfiable, 135~unsatisfiable, and 130~instances of unknown result (cf.~Table~\ref{tab:final}).

@@

\subsection{Planning Instances}
Classical planning is the problem of finding a sequence of actions -- a plan --
that transforms the world from some initial state to a goal state. In 1992 Kautz
and Selman \cite{Kautz1992} proposed to encode planning as satisfiability,
constituting one of the hallmark early adoptions of SAT solving to solve
real-world problems. In their encoding the problem of finding a plan of length
$i$ (\textit{i.e.,} the \emph{makespan}) is translated into a Boolean formula
$F_i$ that is satisfiable if a plan of length $i$ or less exists.
Their encoding is called sequential, whereas parallel encodings allow the execution
of multiple actions in one step~\cite{Rintanen2006, Rintanen2007, Balyo2013}.
Finding the minimal makespan $i$ for which $F_i$ is satisfiable is important for
SAT-based planning in general and the generation of this benchmark set in
particular.
% The minimal makespan depends on the planning task and the used
% encoding.
The hardest formulas that a SAT-based planner has to solve are usually
the last unsatisfiable $F_i$ before the next higher makespan $i$ is
satisfiable~\cite{Rintanen2006}.
% Therefore we preferably pick the last
% unsatisfiable makespan for each planning task to generate unsatisfiable
% instances.

The encodings were generated by two SAT-based planners
\emph{Madagascar}~\cite{Madagascar14} and \emph{Pasar}~\cite{Pasar19}. We used
Madagascar both in its default configuration to generate a parallel encoding
based on $\exists$-step plans and to generate a sequential encoding. Pasar uses
the \emph{grounding routine} deployed by the well-known planner \emph{Fast
  Downward}~\cite{FastDownward06} to translate planning tasks into a different
formalism and then encodes it to SAT using a parallel encoding.
% For planning tasks where this makespan cannot be determined with
% available computational resources, we use a \emph{sequential} encoding, where
% the minimal makespan equals the length of the shortest valid plan. Together with
% known bounds\footnote{Upper and lower bounds on plan length are available for
% 	some planning tasks from the \emph{Optimal track} that have \emph{unit cost}
% 	actions.} on the optimal plan length
%   we can generate satisfiable and unsatisfiable SAT formulas
%   for hard planning problems.
%\textcolor{red}{global note: should try to be consistent with present/past tense...}
The classical planning benchmarks were selected from the \emph{Satisfying} and
\emph{Optimal} tracks of the \emph{International Planning Competitions} 2014
\footnote{\url{https://helios.hud.ac.uk/scommv/IPC-14/repository/benchmarksV1.1.zip}}
and 2018 \footnote{\url{https://bitbucket.org/ipc2018-classical/domains}}. We
only selected planning domains with \emph{unit cost} and eliminated those that
% are designed to be challenging for the grounding routine and
take more than 100 GB of memory to encode into SAT. We ran both Pasar and
Madagascars $\exists$-configuration with a timeout of 3 hours on the
remaining instances to find the minimal makespans. For each planning task where
this process did not timeout, we generated a pair of satisfiable and
unsatisfiable SAT instances. A lot of instances from this set were not used
as \solver{Minisat} could solve them in under $10$ minutes.
% From those we discarded the SAT formulas that \solver{Minisat} can solve in
% under $10$ minutes.
We augmented the remaining domains with the last unsatisfiable formulas
generated for planning task where the minimal makespan could not be found.
To generate the missing benchmarks we use a \emph{sequential} encoding
% , where the minimal makespan equals the length of the shortest valid plan
together with known bounds\footnote{Upper and lower bounds on plan length are available for
	some planning tasks from the \emph{Optimal track} that have \emph{unit cost}
	actions.} on the optimal plan length.

In addition to the classical planning problems, we also included SAT formulas
generated by \emph{Tree-REX} \cite{TreeRex19}, a planner for \emph{Hierarchical
	Task-networks}. In HTN planning the planner is provided with additional domain
knowledge besides the problem description. The HTN benchmarks were provided by
the solver author.

% - 86 out of 100 biggest (number of clauses) are planning
% - 109 / 200 in the planning track are not solved by any solver
% - 186 / 600 in total
% - all 200 above 90\% binary clauses and half have above 99%
The instances of the Planning track are large compared to the Main track
instances. Using the number of clauses as a metric: out of the $100$ biggest
instances of both tracks
% of the main-track
$86$ belong to the Planning track. The large size can mainly be attributed to
binary clauses. On average, more than $98$\% of the clauses are binary for
planning instances. The average for the Main track instances is below $60$\%.

@@

The instances of the Planning track are large compared to the Main track
instances. Using the number of clauses as a metric: out of the $100$ biggest
instances of both tracks
% of the main-track
$86$ belong to the Planning track. The size can be attributed to
binary clauses. On average, more than $98$\% are binary for
planning instances. The average for the Main track instances is below $60$\%.

>>>>

The success of SAT would not be possible without the persistent efforts of the SAT community to further improve the performance and robustness
of SAT solvers. The SAT Competition series, with a history dating back to the early 90s, aims to support and provide further incentives 
for maintaining this progress. Organized yearly as an international open event, SAT Competitions (and their variants
in the forms of SAT Races and a SAT Challenge)~\cite{DBLP:journals/amai/SimonBH05,DBLP:conf/sat/BerreS03,DBLP:conf/sat/BerreS04,DBLP:journals/aim/JarvisaloBRS12,DBLP:journals/ai/BalintBJS15,DBLP:conf/aaai/BalyoHJ17,DBLP:journals/jsat/HeuleJS19}
 have a consistent track record in receiving tens of solver submissions yearly, submitted
by the community at large for obtaining a snapshot of the current state-of-the-art in practical SAT solving. Alongside participating solvers,
the competition invites through open calls submissions of benchmark instances representing, in particular, new interesting applications 
scenarios of SAT solvers. Indeed, in addition to evaluating recently developed solvers, an important aspect of the SAT competition series is
to collect on a yearly basis new benchmark sets, consisting of instances from various different application settings, which together with benchmark sets from previous years constitute a standard dataset for use in research papers and SAT solver development.

@@

However, solvers in No-Limits still competed against all other solvers submitted to the Main Track.
Thus, to deserve a mention, a No-Limits solver would need to rank among the best-performing solvers among the Main Track participants.
In 2020, the top ranked solvers in the No-Limits track were the same as in the Main track. 
This also indicates the stability of results under the exclusion of old benchmark instances.

Complementing the generality advocated by the standard SAT Competition tracks, in which solvers are evaluated
on a set of benchmarks including instances from various types of different problem domains, for 2020 
the organizers aimed to experiment with the potential of a more application-specific track,
each year highlighting a different problem domain where the SAT solving technology helps to advance the state of the art.
In 2020, the Planning track represented the first trial instantiation of this idea.
The focus of this track was specifically on efficiently solving
instances arising from the domain of SAT-based automated planning.\textcolor{red}{Add references} 
Automated planning was chosen as the target problem domain of this first instantiation of the domain-specific tracks
due to its centrality as one of the first breakthrough applications of SAT solvers.
To this end, solvers participating in the Planning
track were evaluated on 200 benchmark instances encoding planning problems.
The same rules for participation as in the Main track applied to the Planning track.
Solvers submitted to the Main track automatically participated in the Planning track.

The traditional Hack track (est. 2009 as \solver{Minisat} Hack track) was organized as a sub-track of the Main track for hacks of \solver{Glucose 3}\textcolor{red}{Add reference}. 
In the past, several advances in SAT solving required only small modifications of an established solver to achieve a considerable contribution. 
Hack tracks encourage participation of such small modifications. 
The limit for being considered a ``hack'' was---somewhat arbitrarily\footnote{The specific threshold for edit distance is not central here; the idea is essentially to only allow relatively small changes to the Glucose
code base, i.e., ``quick hacks'' to Glucose.}---set to 1000 non-space character edit distance from the sources of \solver{Glucose 3}. 
Unfortunately, in 2020 there were not enough participants in this sub-track and so we do not report on it in the results section.

We evaluated all 64 solver submissions (including different configurations of specific solvers) 
to the Main track. Out of the 64 solvers, eight were explicitly submitted to the No-Limits track. Four solvers
 were demoted to the No-Limits track due to outputting invalid unsatisfiability proof certificates. 
Six solvers were disqualified due to outputting truth assignments which did not satisfy the corresponding benchmark instance.
This left us with 46 configurations of 22 solvers, including % only 
one Glucose hack.

@@

In the Incremental Library track the underlying idea is to mimic scenarios
where a SAT solver is used as a back-end solver in a more complex tool
(typically solving a harder problem than SAT) and is called multiple times before 
the enclosing tool reaches its final state. 
``Incremental'' here refers to the idea that the individual calls to the SAT solver are not independent, but may share a common subset of the input clauses or differ in the presence of additional unit clause assumptions~\cite{Nadel:2014:Incremental,Fazekas:2019:IncrementalInprocessing}.\textcolor{red}{Should we not cite the 2004 temporal induction paper by Niklases here?}
Examples for applications of incremental SAT solving are counterexample-guided eastraction refinement (CEGAR) based approaches, e.g., for Bounded Model Checking~\cite{Strichman:2005:ARBMC}, SAT-based planning~\cite{Balyo:2017:IncrPlanning},
multi-agent path finding~\cite{DBLP:conf/ijcai/Surynek19}, and satisfiability modulo theories (SMT) solvers~\cite{Brummayer:2009:Boolector}.

Instead of using or extending the DIMACS input format, in the the Incremental Library track
a general incremental interface called IPASIR (Re-entrant Incremental Solver API) 
is employed~\cite{Balyo:2015:SATRace}. The idea is that we actually run the 
enclosing tool on its own benchmark and communicate with the competing SAT solver 
through this interface API. 
%
SAT solvers submitted to this track need hence implement the require interface.
Furthermore, it should be noted that the solutions output by a solver may, in general, influence
the forthcoming invocations of the solver.
% E.g. Smaller unsat cores may allow the enclosing tool to converge using fewer SAT calls!

@@

The Parallel track evaluates the runtime performance of SAT solvers making use of multiple processor cores
in terms of wall-clock time.
The benchmarks %and the timeout (5000 seconds) 
are the same as in the Main track. 
%\textcolor{red}{repetitive? also mentioned below?}. 
%Each solver in this track was run on a m4.16xlarge machine from Amazon Web Services (AWS).
%These machines have 64 virtual cores and 256GB of memory. 
In contrast to the Main track, proof logging for unsatisfiable instances is not required in the Parallel track.\footnote{
Although this would, of course, be desirable, for the same reasons as in the Main track, 
currently there are no good solutions known for efficient proof logging of parallel solvers.}

@@

For data-driven selection of benchmark instances, we used GBD~Tools\footnote{\url{See https://pypi.org/project/gbd-tools/}} which facilitates querying for instances with desired properties, e.g., by instance author, family, result or solver runtime~\cite{Iser:2018:GBD}. 
We also use GBD~Tools for distributing benchmark instances and their attributes to the general public.\footnote{See \url{https://gbd.iti.kit.edu}}

@@

The ``Bring Your Own Benchmarks'' (BYOB) rule, first  established in SAT Competition 2017~\cite{SC2017}, was again followed
in 2020.
By this rule, solver authors are required to submit 20 ``new'' benchmark instances to accompany a solver submission in order to participate in the competition. 
New refers to the fact that instances contained in benchmark sets from earlier SAT Competitions are not allowed.
Furthermore, a least 10 of the required 20 instances are required to be ``interesting'', interpreted in loose terms by
 the standard \solver{Minisat} SAT solving needing at least one minute of runtime (on typical computing hardware) to solve an instance.
It should be noted that new benchmarks could be submitted to the competition without needing to submit a solver.
As a results, as detailed in Table~\ref{tab:families}, 27 authors contributed a set of 1260 new benchmark instances from a wide range of different instance families.

We decided to include a total of 
300 new benchmarks and a further 100 benchmarks from previous SAT Competitions to the main benchmark set of the 2020 competition.
Key aims of benchmark selection is to ensure that (i) the benchmark set includes enough many relatively hard-to-solve instances
in order to differentiate the overall runtime performances of the competing solvers (without actually running the competing solvers during benchmark selection); 
(ii)~the number of benchmarks included in the benchmark set from different problem domains 
is balanced across the problem domains, and that (iii) the benchmark set is also balanced in terms of the number of unsatisfiable and satisfiable instances included in the set.

To compile the set of 300 new instances, we first applied a hardness criterion by filtering out all instances solved by \solver{Minisat} in less than 10 minutes.\footnote{We note that
the limit of 10 minutes is again somewhat arbitrary.
This 10-min runtime hardness filter essentially aims to make sure that enough instances are included in the final benchmark set which allow for distinguishing in terms of relative performance between the best-performing competing solvers.
Unfortunately this time the 10-min limit enforced  was much greater that the requirements imposed for  ``interesting'' benchmark instances
by the BYOB rule. It could be more sensible to impose the some 10-min limited also for an instance being ``interesting''. However, this would require more efforts
at least in terms of computation times from the solver authors in order to construct a set of 20 interesting new benchmarks required by the BYOB rule.}

From the resulting 1012 instances, in order to obtain a balanced benchmark set, 
we randomly selected $k$~instances per author using the value  $k$ which ensured that  the resulting set contains at least 300 instances.
This \emph{rule-based randomization} procedure is detailed as
Algorithm~\ref{algo:select}. Specifically,
we randomly selected 7~satisfiable and 7~unsatisfiable instances per author (Lines~3 and~4) and added instances of yet unknown result if this did not yield a total of 14~instances (Lines~5--7). 
Of the such obtained 308~instances, we randomly removed 8~satisfiable instances, yielding a total of 114~satisfiable, 78~unsatisfiable and 108~instances of unknown satisfiability
status.

We augmented the then obtained set of 300 new benchmarks with 100 instances from previous SAT competitions as follows. 
In order to further balance the number of satisfiable and unsatisfiable instances in the new benchmark set, we randomly selected 21~satisfiable, 57~unsatisfiable and 22~unknown instances. 
With additional constraints, we made sure not to select instances from benchmark 
families which are already represented in the set of~300 new instances (cf. Table~\ref{tab:families}). 
We also excluded random, agile and planning instances (due to the Planning track). 
The final main benchmark set  contains 135~satisfiable, 135~unsatisfiable, and 130~instances of ``unknown'' status (cf.~Table~\ref{tab:final}).

@@

\subsection{Planning Instances}
Classical planning is the problem of finding a sequence of actions---a plan---that 
transforms the world from some initial state to a goal state. In 1992 Kautz
and Selman \cite{Kautz1992} proposed to encode planning as satisfiability,
constituting one of the hallmark early adoptions of SAT solving to solve
real-world problems. In their encoding the problem of finding a plan of length
$i$ (\textit{i.e.,} the \emph{makespan}) is translated into a Boolean formula
$F_i$ that is satisfiable if a plan of length $i$ or less exists.
Their encoding is called sequential, whereas parallel encodings allow the execution
of multiple actions in one step~\cite{Rintanen2006, Rintanen2007, Balyo2013}.
Finding the smallest makespan $i$ for which $F_i$ is satisfiable is important for
SAT-based planning in general and the generation of this benchmark set in
particular.
% The minimal makespan depends on the planning task and the used
% encoding.
The hardest formulas that a SAT-based planner has to solve are usually
the last unsatisfiable $F_i$ before the next higher makespan $i+1$ is
satisfiable~\cite{Rintanen2006}.
% Therefore we preferably pick the last
% unsatisfiable makespan for each planning task to generate unsatisfiable
% instances.

For the Planning track,
the benchmark instances were generated using two SAT-based planners
\emph{Madagascar}~\cite{Madagascar14} and \emph{Pasar}~\cite{Pasar19}. We used
Madagascar both in its default configuration to generate a parallel encoding
based on $\exists$-step plans and to generate a sequential encoding. Pasar uses
the \emph{grounding routine} deployed by the well-known planner \emph{Fast
  Downward}~\cite{FastDownward06} to translate planning tasks into a different
formalism and then encodes it to SAT using a parallel encoding.
% For planning tasks where this makespan cannot be determined with
% available computational resources, we use a \emph{sequential} encoding, where
% the minimal makespan equals the length of the shortest valid plan. Together with
% known bounds\footnote{Upper and lower bounds on plan length are available for
% 	some planning tasks from the \emph{Optimal track} that have \emph{unit cost}
% 	actions.} on the optimal plan length
%   we can generate satisfiable and unsatisfiable SAT formulas
%   for hard planning problems.
%\textcolor{red}{global note: should try to be consistent with present/past tense...}
The classical planning benchmarks were selected from the \emph{Satisfying} and
\emph{Optimal} tracks of the \emph{International Planning Competitions} 2014
\footnote{\url{https://helios.hud.ac.uk/scommv/IPC-14/repository/benchmarksV1.1.zip}}
and 2018 \footnote{\url{https://bitbucket.org/ipc2018-classical/domains}}. We
only selected planning domains with \emph{unit cost} and eliminated those that
% are designed to be challenging for the grounding routine and
take more than 100 GB of memory to encode into SAT. We ran both Pasar and
Madagascars $\exists$-configuration with a timeout of 3 hours on the
remaining instances to find the minimal makespans. For each planning task where
this process did not timeout, we generated a pair of satisfiable and
unsatisfiable SAT instances. A significant number of instances from this set were not used
as \solver{Minisat} could solve them in under $10$ minutes.
% From those we discarded the SAT formulas that \solver{Minisat} can solve in
% under $10$ minutes.
We augmented the remaining domains with the last unsatisfiable formulas
generated for planning task where the minimal makespan could not be found.
To generate the missing benchmarks we use a \emph{sequential} encoding
% , where the minimal makespan equals the length of the shortest valid plan
together with known bounds\footnote{Upper and lower bounds on plan length are available for
	some planning tasks from the \emph{Optimal track} that have \emph{unit cost}
	actions.} on the optimal plan length.

In addition to the classical planning problems, we also included SAT instances
generated by \emph{Tree-REX} \cite{TreeRex19}, a planner for \emph{Hierarchical
	Task-networks} (HTN). In HTN planning the planner is provided with additional domain
knowledge besides the problem description. The HTN benchmarks were provided by
the solver author.

% - 86 out of 100 biggest (number of clauses) are planning
% - 109 / 200 in the planning track are not solved by any solver
% - 186 / 600 in total
% - all 200 above 90\% binary clauses and half have above 99%
The instances of the Planning track are large in size compared to the Main track
instances. Using the number of clauses as a metric, out of the $100$ largest
instances across both tracks,
% of the main-track
$86$ belong to the Planning track benchmark set. The large size of Planning track instances
can mainly be attributed to large numbers of
binary clauses that SAT encodings of planning problems naturally produce. 
On average, more than $98$\% of the clauses are binary for
planning instances. The average for the Main track instances is below $60$\%.

@@
